### 2024-05-28 workflow/notebooks/eda.Rmd
added notebooks, FTPY and some code improvements

---
title: "Structural analysis"
# knit: (function(input_file, encoding) {
#   rmarkdown::render(input_file,
date: "`r format(Sys.time(), '%d-%b-%Y')`"
author: "Giacomo Mutti"
output:
    rmdformats::html_clean:
        number_sections: false
        code_folding: hide
        self_contained: true
        fig_caption: true
        # thumbnails: false
        # lightbox: false
        gallery: true
        toc_depth: 3
        highlight: kate
---

```{r setup, include=FALSE, dev="CairoPNG"}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, cache.lazy = FALSE,
                      crop = TRUE, dev="CairoPNG")
knitr::knit_hooks$set(crop = knitr::hook_pdfcrop)
knitr::opts_knit$set(root.dir = "../../")
```

```{r libraries, message=FALSE, warning=FALSE, cache=FALSE}
library(tidyverse)
library(patchwork)
library(ggtree)
library(cowplot)
library(glue)
source("workflow/scripts/functions.R")
theme_set(theme_classic())

models <- names(palettes_model)

params <- yaml::read_yaml("config/params.yaml")
params_dataset <- yaml::read_yaml("config/Hsapopi_test.yaml")

outdir <- glue(params$outdir, params_dataset$dataset)
```

```{r read_stuff, cache=FALSE}
seeds <- readLines(glue(outdir, "/ids/UP000005640_common.ids"))

tax <- read_delim(params_dataset$taxons_file, show_col_types=F)
taxmap <- read_delim(glue(outdir, "/db/taxidmap"), 
                     delim = "\t", col_names = c("label", "Tax_ID"),
                     show_col_types = FALSE) %>% 
  left_join(tax, by = "Tax_ID") %>% 
  dplyr::select(label, Proteome_ID, mnemo)

taxidmap <- read_delim(glue(outdir,"/db/taxidmap"), 
                       col_names = c("target", "Tax_ID"), show_col_types=F)

sptree <- read.tree(params_dataset$species_tree_labels)
translate <- deframe(distinct(dplyr::select(taxmap, Proteome_ID, mnemo)))
# sptree$tip.label <- names(translate)[match(translate, sptree$tip.label)]

species_dist_df <- tibble()
for (sp in sptree$tip.label){
  if (sp=="Hsap") {next}
  dist <- castor::get_all_distances_to_tip(sptree, "Hsap")[MRCA(sptree, "Hsap", sp)]
  mrca <- c(sptree$tip.label, sptree$node.label)[MRCA(sptree, "Hsap", sp)]
  species_dist_df <- bind_rows(species_dist_df, tibble(mnemo=sp, value=dist, mrca=mrca))
}

# read input table
lvls_tax <- rev(c(unique(tax$Clade), "archaea", "bacteria"))

table_columns <- c('Proteome_ID', 'Tax_ID', 'count1', 'count2', 'count3', 
                   'genome', 'source', 'species', 'mnemo')

table <- read_delim(params_dataset$input_table_file, show_col_types=F, 
                    delim = "\t", col_names = table_columns) %>% 
  left_join(tax, by = "Tax_ID") %>% 
  mutate(Clade = ifelse(is.na(Clade), mnemo, Clade),
         Clade = factor(Clade, levels = lvls_tax))
```

# Homology

```{r read_self_blast}
blast_self <- read_delim(glue(outdir, "/homology/allvall/",
                              params_dataset$seed, "_", params_dataset$seed, "_blast.tsv"),
                         show_col_types=FALSE,
                         col_names = blast_columns) %>% 
  filter(query %in% seeds) %>%
  filter(query==target) %>% 
  mutate(blast_selfbit=bitscore) %>% 
  select(query, blast_selfbit)

fs_self <- read_delim(glue(outdir, "/homology/allvall/",
                              params_dataset$seed, "_", params_dataset$seed, "_fs.tsv"), 
                      show_col_types = FALSE,
                         col_names = blast_columns) %>% 
  filter(query %in% seeds) %>%
  filter(query==target) %>% 
  mutate(fs_selfbit=bitscore) %>% 
  select(query, fs_selfbit)

self <- full_join(blast_self, fs_self, by="query")
```

```{r read_fs}
fs <- read_delim(glue(outdir, "/homology/", params_dataset$seed, "_fs.tsv"), 
                 col_names = fs_columns, show_col_types = FALSE) %>% 
  filter(query %in% seeds) %>%
  mutate(evalue = ifelse(evalue==0, 1e-180, evalue)) %>%
  left_join(taxidmap, by = "target")

fs_brhs <- read_delim(glue(outdir, "/homology/", params_dataset$seed, "_fs_brh.tsv"),
                     col_names = c("query", "target"), show_col_types = FALSE) %>% 
  filter(query %in% seeds) %>%
  mutate(method="fs_brh") 
```

```{r read_blast}
blast <- read_delim(glue(outdir, "/homology/", params_dataset$seed, "_blast.tsv"), 
                    col_names = blast_columns, show_col_types = FALSE) %>% 
  filter(query %in% seeds) %>%
  mutate(evalue = ifelse(evalue==0, 1e-180, evalue))

blast_brhs <- read_delim(glue(outdir, "/homology/", params_dataset$seed, "_blast_brh.tsv"),
                     col_names = c("query", "target"), show_col_types = FALSE) %>% 
  filter(query %in% seeds) %>%
  mutate(method="blast_brh")
```

```{r merge_homo}
df <- full_join(blast, fs,
                by = c("query", "target"), suffix=c("_blast", "_fs")) %>%
    filter(query!=target) %>% 
  left_join(self, by="query") %>%
  mutate(pident_fs = 100*pident_fs, 
         qcov_fs = 100*qcov_fs,
         singleton = case_when(is.na(evalue_fs) ~ "only_blast",
                               is.na(evalue_blast) ~ "only_fs", 
                               .default = "common"),
         blast_norm = bitscore_blast/blast_selfbit, fs_norm = bitscore_fs/fs_selfbit)

brhs <- rbind(blast_brhs, fs_brhs) %>% 
    group_by(query, target, method) %>%
    mutate(n=1) %>%
    pivot_wider(names_from = method, values_fill = 0, values_from = n) %>%
    mutate(common = ifelse(fs_brh+blast_brh==2, 1, 0))

# grouped_brhs <- brhs %>%
#   group_by(query) %>%
#   summarise(fs_brh = sum(fs_brh), blast_brh=sum(blast_brh), common_brh=sum(common))
```
## Blast vs Foldseek

There are these number of hits, divided by singleton class:

```{r eda_homo}
df %>% count(singleton)
```

We can explore different properties of blast and foldseek search divided by singleton class.

<!-- ```{r blast_vs_fs, echo = FALSE, message=FALSE, fig.align='center', fig.cap='Some cool caption', out.width='\\linewidth', fig.pos='H'} -->
<!-- knitr::include_graphics(glue(outdir, "/plots/", params_dataset$seed, "_homology.pdf")) -->
<!-- ``` -->

<center>
![*Blast vs foldseek properties*](../../draft/figures/UP000005640_homology.png)
</center>

## BRHs

We can explore the abundance of BRH hits for every seed protein across the different species.

```{r brhs, fig.width=12}
brhs %>% 
  left_join(select(taxmap, -Proteome_ID), by=c("target"="label")) %>% 
  mutate(mnemo=factor(mnemo, levels=sptree$tip.label)) %>% 
  # pivot_longer(cols = c(blast_brh, fs_brh, common)) %>% 
  filter(mnemo!="Hsap") %>% 
  pivot_longer(cols = c(blast_brh, fs_brh, common)) %>% 
  filter(value!=0) %>% 
  ggplot(aes(mnemo, fill=name)) + 
  geom_bar(position = "dodge") + 
  scale_fill_manual(values = palette_brh) + 
  scale_y_continuous(expand = expansion(0,0))
# select(df, query, target, singleton) %>% 
#   left_join(brhs) %>% 
#   group_by(singleton) %>%
#   summarise(#prop_brh = sum(!is.na(common))/n()*100,
#             prop_blast_brh = sum(blast_brh, na.rm = T)/n()*100, 
#             prop_fs_brh = sum(fs_brh, na.rm = T)/n()*100, 
#             prop_common_brh = sum(common, na.rm = T)/n()*100) %>% 
#   pivot_longer(!singleton) %>% 
#   ggplot(aes(name, value, color=singleton)) + 
#   geom_point(size=3) + 
#   scale_color_manual(values=palette_singleton) + 
#   labs(x="Type of BRH", y="% of BRH")
```

## Singletons structure quality

```{r lddt_singletons, fig.width=10, fig.height=5}
lddt_df <- read_delim(paste0(params$structure_dir, "/", tax$Proteome_ID, "/", tax$Proteome_ID, "_mean_plddt.tsv"),
           col_names = c("target", "mean_lddt"), show_col_types = FALSE)

df %>% 
  select(target, singleton) %>% 
  mutate(target=gsub("AF-|-F1", "", target)) %>% 
  left_join(lddt_df, by="target") %>% 
  ggplot(aes(mean_lddt, fill=singleton)) + 
  geom_density(alpha=.5) +
  scale_y_continuous(expand = expansion(0.01,0)) +
  scale_x_continuous(expand = expansion(0.01,0)) +
  scale_fill_manual(values = palette_singleton) 
```

We can see the meand plddt for singleton classes and common are generally better imputed proteins than the two singletons. Expectedly fs suffers more than blast.

We can also see more in detail some blast and foldseek features and how they differ.

<center>
![*Foldseek and blast properties divided by singleton class*](../../draft/figures/UP000005640_singletons.png)
</center>

# Post filtering blast

```{r filter_homo}
aln_seeds <- paste0("AF-",readLines(glue(outdir, "/ids/",params_dataset$seed,"_aln.ids")), "-F1")
blast_filtered <- filter(blast, 
                         query %in% aln_seeds,
                         length/qlen*100>params$coverage, 
                         length/slen*100>params$coverage,
                         evalue<as.numeric(params$eval_both)) %>% 
  group_by(query) %>% 
  slice_head(n = params$max_seqs) %>% 
  ungroup()

fs_filtered <- filter(fs, 
                      query %in% aln_seeds,
                      qcov*100>params$coverage, 
                      tcov*100>params$coverage, 
                      evalue<as.numeric(params$eval_both)) %>% 
  group_by(query) %>% 
  slice_head(n = params$max_seqs) %>% 
  ungroup()
```

We filter blast and foldseek by `r params$coverage` query and target coverage and by `r params$eval_both` min evalue for our `r length(seeds)`. Only `r length(aln_seeds)` will go in the phylogenetic pipeline as they have more than 4 common hits. 
This results in `r nrow(blast_filtered)` blast rows and `r nrow(fs_filtered)` fs rows.

## Originations

```{r originations}
blast_originations <- blast_filtered %>% 
  select(query, target) %>% 
  left_join(taxmap, by=c("target"="label")) %>% 
  group_by(query) %>% 
  filter(mnemo!="Hsap") %>% 
  left_join(species_dist_df, by="mnemo") %>% 
  mutate(mrca=factor(mrca, levels=unique(species_dist_df$mrca), ordered = TRUE)) %>% 
  summarise(mrca=max(mrca)) %>% 
  group_by(mrca) %>% 
  count() %>% 
  mutate(method="blast")

fs_originations <- fs_filtered %>% 
  select(query, target) %>% 
  left_join(taxmap, by=c("target"="label")) %>% 
  group_by(query) %>% 
  filter(mnemo!="Hsap") %>% 
  left_join(species_dist_df, by="mnemo") %>% 
  mutate(mrca=factor(mrca, levels=unique(species_dist_df$mrca), ordered = TRUE)) %>% 
  summarise(mrca=max(mrca)) %>% 
  group_by(mrca) %>% 
  count() %>% 
  mutate(method="fs")

originations <- rbind(blast_originations, fs_originations)

originations %>% 
  ggplot(aes(n, fct_rev(mrca), fill=method)) + 
  geom_bar(stat = "identity", position = "dodge") + 
  scale_fill_manual(values = palettes_method) + 
  scale_x_continuous(expand = expansion(0,0)) + 
  labs(y="MRCA")
```

We can see where is the "taxonomic mrca" (the most distant species) for every seed. Interestingly, foldssek is more than blast just in the two last nodes.

## Duplications

```{r duplications}
duplications_mrca <- read_delim(glue(outdir, "/reco/", params_dataset$seed,"_mrca.tsv"),
                                show_col_types = FALSE) %>% 
  separate(id, c("gene", "target", "alphabet", "model")) %>% 
  select(-alphabet) %>% 
  group_by(target, model, mrca) %>% 
  summarise(n=sum(n)) %>% 
  filter(model %in% c("LG", "FTPY")) %>% 
  mutate(mrca=factor(mrca, levels=fortify(sptree) %>% arrange(y) %>% pull(label), ordered = TRUE))

duplications_mrca %>% 
  filter(target %in% c("blast", "fs")) %>%
  ggplot(aes(n, mrca, fill=target)) + 
  geom_bar(position = "dodge", stat="identity") +
  facet_grid(~model) + 
  scale_x_continuous(expand = expansion(0,0)) +
  scale_fill_manual(values=palettes_method)

# fortify(sptree) %>% 
#   inner_join(duplications_mrca, by=c("label"="mrca")) %>% 
#   ggtree() + 
#   geom_point(aes(size=n)) + 
#   facet_grid(target~model)
```

We can also explore where the duplications called by RangerDTL are mapped in the species tree. For this we are only interested in *LG* and *FTPY* to reduce complexity.

# Trees

```{r read_trees}
df_trees <- read_delim(glue(outdir, "/trees/", params_dataset$seed, "_unrooted_trees.txt"),
                       show_col_types = FALSE,
                       col_names = c("gene", "target", "alphabet", "model", "tree")) %>% 
  mutate(model=factor(model, levels=models))
```

## Reconciliation

<center>
![*all 6 model results*](../../draft/figures/UP000005640_discordance.png)
</center>

<center>
![*Astral-PRO results*](../../draft/figures/UP000005640_astral_pro.png)
</center>

<center>
![*Distance 2 seed by singleton class*](../../draft/figures/UP000005640_distance.png)
</center>

## RF distance + varr2t

<center>
![*Various Tree analysis*](../../draft/figures/UP000005640_trees.png)
</center>

## Model fit

How many times is each model better?

```{r prop_better, fig.width=10}
reco <- read_delim(glue(outdir, "/reco/", params_dataset$seed, "_DTL.tsv"), show_col_types = FALSE) %>%
  mutate(model=factor(model, levels=models))

scores <- read_delim(glue(outdir, "/reco/", params_dataset$seed, "_scores.tsv"), show_col_types = FALSE) %>%
  left_join(reco) %>%
  mutate(model=factor(model, levels=models))

reco_prop_plot <- scores %>%
  group_by(id, targets) %>%
  mutate(norm_reco_score=(dups+losses)/n_tips) %>%
  slice_min(norm_reco_score, with_ties = TRUE) %>%
  mutate(best_score=1/n()) %>% 
  ggplot(aes(targets, best_score, fill=model)) +
  geom_bar(stat = "identity") +
  labs(subtitle = "How many times each model has the lowest D+L/tips?") +
  scale_fill_manual(values = palettes_model)

tcs_prop_plot <- scores %>%
  group_by(id, targets) %>%
  slice_max(score, with_ties = TRUE) %>%
  mutate(best_score=1/n()) %>% 
  ggplot(aes(targets, best_score, fill=model)) +
  geom_bar(stat = "identity") +
  labs(subtitle = "How many times each model has the highest TCS score?") +
  scale_fill_manual(values = palettes_model)

reco_prop_plot + tcs_prop_plot +  plot_layout(guides = 'collect')
```

## Distance between target classes

We can see the average normalized topological distance between the different groups in LG and FTPY.

```{r targets_topodist, fig.width=10, fig.height=5}
df_trees_dist <- filter(df_trees, model %in% c("LG", "FTPY"), target=="union")
ts_dist <- read.tree(text = df_trees_dist$tree)
names(ts_dist) <- paste(df_trees_dist$gene, df_trees_dist$target, df_trees_dist$alphabet, df_trees_dist$model, sep = "_")

df_distance <- tibble()

for (idx in 1:length(ts_dist)) {
    seed <- paste0("AF-",df_trees_dist$gene[idx],"-F1")
    tmp_df <- select(filter(df, query==seed), target, singleton)

    tree_dist <- as.matrix(adephylo::distTips(ts_dist[[idx]], method = "nNodes")) %>% 
      as.data.frame() %>% 
      rownames_to_column() %>% 
      pivot_longer(!rowname) %>% 
      filter(rowname>name) %>%
      mutate(value=value/max(value)) %>% 
      left_join(tmp_df, by = c("rowname"="target")) %>% 
      left_join(tmp_df, by = c("name"="target")) %>% 
      mutate(a=pmin(singleton.x, singleton.y),
             b=pmax(singleton.x, singleton.y)) %>% 
      group_by(a, b) %>%
      filter(!is.na(a), !is.na(b)) %>%
      summarise(mean_dist=mean(value), n=n(), .groups = "keep") %>%
      mutate(gene=seed, model=df_trees_dist$model[idx])

  df_distance <- bind_rows(df_distance, tree_dist)
}

df_distance %>% 
  group_by(a,b,model) %>% 
  summarise(wmean=weighted.mean(mean_dist,n)) %>% 
  ggplot(aes(a,b, fill=wmean)) + 
  geom_tile() +
  geom_text(aes(label=round(wmean,2))) +
  facet_grid(~model) +
  scale_fill_gradientn(colors=wespal) +
  coord_cartesian(expand = F) +
  theme(axis.title = element_blank(),
        legend.position = "none")
```

# CATH/SCOPe

# Runtimes

<center>
![*Various runtimes statistics*](../../draft/figures/UP000005640_runtime.png)
</center>

# TODOs

* CATH/SCOPe
* Benchmark: fix runtimes!
* Fix FTPY branch support
* Foldtree smk vs Foldtree python: DONE
* duplication in species tree: DONE