### 2024-05-28 workflow/notebooks/explore_examples.Rmd

---
title: "Structural vs Sequence Dashboard"
runtime: shiny
output:
  flexdashboard::flex_dashboard:
    theme:
      bg: "#FFFFFF"
      fg: "#00a276"
      primary: "#00a276"
      base_font:
        google: Roboto
      code_font:
        google: JetBrains Mono
    orientation: columns
    vertical_layout: fill
---

```{r setup, include=FALSE}
library(flexdashboard)
# Install thematic and un-comment for themed static plots (i.e., ggplot2)
# thematic::thematic_rmd()
```

## Column {.tabset data-width="650"}

```{r libraries, message=FALSE, warning=FALSE}
suppressPackageStartupMessages(library(ggtree))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(ggmsa))
library(patchwork)
library(shiny)

source("../scripts/functions.R")
theme_set(theme_classic())
```

```{r input_files}
models <- c("QT", "FT", "LG", "GTR", "3Di")

ids_input <- "../../results/test/ids/UP000005640_examples.ids"
trees <- "../../results/test/trees/UP000005640_unrooted_trees.txt"
taxidmap <- "../../results/test/db/taxidmap"
meta <- "../../data/meta/Hsap_taxon.csv"

seeds <- readLines(ids_input)
seed_sp <- gsub("_examples.ids", "", basename(ids_input))
seed_dir <- gsub("trees", "seeds", dirname(trees))

taxmap <- read_delim(taxidmap, 
                     delim = "\t", col_names = c("label", "Tax_ID")) %>% 
  left_join(read_delim(meta)) %>% 
  dplyr::select(label, Proteome_ID, mnemo)

fls <- list.files(paste0("~/mn5/projects/projects/structural_phylome/results/test/seeds/", seed_sp, "/", seeds),
                  full.names = TRUE)
ids <- fls[grepl("(blast|fs).*ids$", fls)]
x <- sapply(ids, readLines, simplify = FALSE)
names(x) <- gsub("\\.ids", "", basename(ids))
alns <- fls[grepl("commo.*clean$", fls)]
```

```{r}
selectInput("seed", label = "seed", choices = seeds)
```

```{r}
venn_input <- reactive({
  x[which(grepl(input$seed, names(x)))]
})

aln_3di_input <- reactive({
  Biostrings::readAAMultipleAlignment(filepath = alns[grepl(paste0(input$seed,".*3Di.alg"), alns)])
})

aln_aa_input <- reactive({
  Biostrings::readAAMultipleAlignment(filepath = alns[grepl(paste0(input$seed,".*aa.alg"), alns)])
})

# alns <- fls[grepl("commo.*clean$", fls)]
#   
# # 3Di aln
# aln_3di <- Biostrings::readAAMultipleAlignment(alns[grepl("3Di.alg", alns)])
# aln_3di_mat <- as.matrix(aln_3di@unmasked)
# pos_df_3di <- tibble(pos=1:ncol(aln_3di_mat), 
#                      cons=nrow(aln_3di_mat) - colSums(aln_3di_mat=="-" | aln_3di_mat=="X"))
# 
# aln_aa <- Biostrings::readAAMultipleAlignment(alns[grepl("aa.alg", alns)])
# aln_aa_mat <- as.matrix(aln_aa@unmasked)
# pos_df_aa <- tibble(pos=1:ncol(aln_aa_mat), 
#                      cons=nrow(aln_aa_mat) - colSums(aln_aa_mat=="-" | aln_aa_mat=="X"))
# 
# trees_files <- fls[grepl("nwk$", fls)]
# 
# trees <- read.tree(text = sapply(trees_files, readLines))
# names(trees) <- gsub(paste0(seed,"_|(3Di|aa)_|.nwk$"), "", basename(trees_files))

# conf_df <- tibble(label=union(x$blast, x$fs)) %>% 
#   left_join(taxmap) %>% 
#   mutate(confidence_file=paste0("results/data/structures/",
#                                 Proteome_ID,
#                                 "/confidence/",
#                                 label,
#                                 "-confidence_v4.json.gz")) %>% 
#   rowwise() %>% 
#   mutate(conf=mean(jsonlite::fromJSON(gzfile(confidence_file))$confidenceScore))
```

### Venn

```{r}
renderPlot(
  venn_input() %>% 
    ggvenn::ggvenn(auto_scale = TRUE)
)
```

### Alns 3di

```{r aln 3di}

renderPlot(
  aln_3di_input() %>% 
  ggmsa(seq_name = FALSE, font = NULL, show.legend = TRUE, border = NA) +
    labs(title = paste(input$seed, "- 3Di common hits cleaned aln")) +
    # geom_line(mapping = aes(x=pos, y=cons), data = pos_df_3di, inherit.aes = F) +
    coord_cartesian() + 
    theme_void() + 
    theme(axis.text.x = element_text())
)

```

### Aln AA

```{r aln_aa}

renderPlot(
  aln_aa_input() %>% 
  ggmsa(seq_name = FALSE, font = NULL, show.legend = TRUE, border = NA) +
    labs(title = paste(input$seed, "- 3Di common hits cleaned aln")) +
    # geom_line(mapping = aes(x=pos, y=cons), data = pos_df_3di, inherit.aes = F) +
    coord_cartesian() + 
    theme_void() + 
    theme(axis.text.x = element_text())
)

```