### 2024-06-04 workflow/notebooks/CATH.Rmd
Added CATH benchmark

---
title: "CATH analysis"
# knit: (function(input_file, encoding) {
#   rmarkdown::render(input_file,
date: "`r format(Sys.time(), '%d-%b-%Y')`"
author: "Giacomo Mutti"
output:
    rmdformats::html_clean:
        number_sections: false
        code_folding: hide
        self_contained: true
        fig_caption: true
        # thumbnails: false
        # lightbox: false
        gallery: true
        toc_depth: 3
        highlight: kate
---

```{r setup, include=FALSE, dev="CairoPNG"}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, cache.lazy = FALSE,
                      crop = TRUE, dev = "CairoPNG")
knitr::knit_hooks$set(crop = knitr::hook_pdfcrop)
knitr::opts_knit$set(root.dir = "../../")
```

```{r libraries, message=FALSE, warning=FALSE, cache=FALSE}
library(tidyverse)
library(ggraph)
library(tidygraph)
library(patchwork)
library(ggtree)
library(cowplot)
library(glue)
source("workflow/scripts/functions.R")
theme_set(theme_classic())

models <- names(palettes_model)

params <- yaml::read_yaml("config/params.yaml")
params_dataset <- yaml::read_yaml("config/Hsapopi_test.yaml")

outdir <- glue(params$outdir, params_dataset$dataset)
```

```{r read_dfs}
CATH_df <- read_delim(list.files(params$cath_dir, full.names = TRUE),
                      delim = "\t", id = "proteome", show_col_types = FALSE) %>%
  # filter(!is.na(Gene3D)) %>% 
  mutate(proteome = gsub("_cath.tsv", "", basename(proteome)),
         n_3d_domains = str_count(Gene3D, ";"),
         n_Pfam_domains = str_count(Pfam, ";")) %>%
  # filter(n_domains==1) %>% 
  mutate(target=paste0("AF-", Entry, "-F1")) %>% 
    select(-proteome, Entry)

nms <- read_delim("test/cath/cath-names.txt", 
                  comment = "#", delim = "    ",
                  col_names = c("Gene3D", "abbr", "name"), 
                  show_col_types = FALSE) %>% 
  select(-abbr) %>% 
  mutate(name=gsub("^:", "", name))

nms_pfam <- read_delim("test/cath/Pfam-A.clans.tsv", 
           col_names = c("Pfam", "clan", "boh", "abbr", "name"), 
           show_col_types = FALSE) %>% 
  select(-name, -clan, -boh)

long_CATH_df <- CATH_df %>% 
  # filter(target %in% unique(blast$query)) %>% 
  mutate(Gene3D=strsplit(Gene3D, ";")) %>% 
  unnest(Gene3D) %>% 
  separate(col = Gene3D, into = c("C", "A", "T", "H"), sep = "\\.", remove = FALSE) %>% 
  mutate(CA = paste(C,A,sep = "."), CAT=paste(C,A,T,sep = ".")) %>% 
  left_join(nms, by=c("C"="Gene3D")) %>%  
  left_join(nms, by=c("CA"="Gene3D"), suffix = c("_C", "_CA")) %>% 
  left_join(nms, by=c("CAT"="Gene3D"), suffix = c("_CA", "_CAT")) %>% 
  left_join(nms, by=c("Gene3D"), suffix = c("_CAT", "_CATH"))

long_CATH_nms <- long_CATH_df %>%
    select(-c(Entry, Pfam, n_Pfam_domains, target, n_3d_domains)) %>% 
    distinct()
```

```{r read_blast}
seeds <- paste0("AF-",readLines(glue(outdir, "/ids/",params_dataset$seed,"_aln.ids")), "-F1")

fs <- read_delim(glue(outdir, "/homology/", params_dataset$seed, "_fs.tsv"), 
                 col_names = fs_columns, show_col_types = FALSE) %>% 
  # filter(query %in% seeds) %>%
  mutate(evalue = ifelse(evalue==0, 1e-180, evalue)) %>% 
  filter(qcov*100>params$coverage,
         tcov*100>params$coverage,
         evalue<as.numeric(params$eval_both)) %>% 
  group_by(query) %>% 
  slice_head(n = params$max_seqs) %>% 
  ungroup()

blast <- read_delim(glue(outdir, "/homology/", params_dataset$seed, "_blast.tsv"), 
                    col_names = blast_columns, show_col_types = FALSE) %>% 
  # filter(query %in% seeds) %>%
  mutate(evalue = ifelse(evalue==0, 1e-180, evalue)) %>% 
  filter(length/qlen*100>params$coverage, 
         length/slen*100>params$coverage,
         evalue<as.numeric(params$eval_both)) %>% 
  group_by(query) %>% 
  slice_head(n = params$max_seqs) %>% 
  ungroup()

df <- full_join(mutate(select(fs, query, target), method="fs"),
                mutate(select(blast, query, target), method="blast"),
          by = c("query", "target"), suffix=c("_fs", "_blast")) %>% 
    filter(query!=target) %>% 
    mutate(singleton = case_when(is.na(method_fs) ~ "only_blast",
                                 is.na(method_blast) ~ "only_fs", 
                                 TRUE ~ "common")) %>% 
  select(-contains("method"))
```

# CATH exploration

First of all we can see more or less how many domains (3d and Pfam) there are in all species?

By analyzing only monodomains we simplify everything and we still consider the majority of the data.

```{r monodomain, fig.width=6, fig.height=5}
CATH_df %>% 
  pivot_longer(c(n_3d_domains, n_Pfam_domains), values_to = "N_domains") %>% 
  mutate(name=str_split_i(name, "_", 2)) %>% 
  ggplot(aes(N_domains, fill=name)) + 
  geom_histogram(binwidth = 1, position = "dodge", color="black") +
  scale_x_continuous(breaks = seq(1,100,1)) +
  scale_y_continuous(expand = expansion(0,0)) + 
  scale_fill_manual(values=palette_domains)
```

From now on we will analyze the results based on the top 150 filtered hits for each query (not only the test seeds).
We can check how many distinct CATHs there are in targets in monodomain queries. Here it shows that foldseek recovers more weird stuff (analogous etc).

```{r monodomain_distinct, fig.width=12, fig.height=8}
plot_CATH <- inner_join(df, CATH_df, by = "target") %>%
  filter(n_3d_domains==1) %>%
  group_by(query, singleton) %>% 
  summarise(n_distinct_CATH = n_distinct(Gene3D)) %>% 
  ggplot(aes(n_distinct_CATH, fill=singleton)) + 
  geom_histogram(binwidth = 1, position = "dodge") +
  # facet_grid(n_domains~., scales = "free_y") +
  scale_x_continuous(breaks = seq(1,100,1)) + 
  scale_y_continuous(expand = expansion(0,0)) + 
  scale_fill_manual(values=palette_singleton)

plot_Pfam <- inner_join(df, CATH_df, by = "target") %>%
  filter(n_Pfam_domains==1) %>% 
  group_by(query, singleton) %>% 
  summarise(n_distinct_Pfam = n_distinct(Pfam)) %>% 
  ggplot(aes(n_distinct_Pfam, fill=singleton)) + 
  geom_histogram(binwidth = 1, position = "dodge") +
  # facet_grid(n_domains~., scales = "free_y") +
  scale_x_continuous(breaks = seq(1,100,1)) + 
  scale_y_continuous(expand = expansion(0,0)) + 
  scale_fill_manual(values=palette_singleton)

(plot_CATH / plot_Pfam) + plot_layout(guides = 'collect')
```

We can also see when the Pfam and CATH disagree in monodomain proteins, this could be explained by the fact the structure is more conserved, therefore less FALSEs where sequence is more prone to this.

```{r combn, fig.width=4, fig.height=8}
df %>% 
  select(-contains("method")) %>%
  left_join(CATH_df, by=c("query"="target")) %>% 
  left_join(CATH_df, by=c("target"), suffix = c("_query", "_target")) %>% 
  filter(query!=target, n_Pfam_domains_query==1, n_3d_domains_query==1,
          n_Pfam_domains_target==1, n_3d_domains_target==1) %>% 
  mutate(is_same_3d=Gene3D_query==Gene3D_target, is_same_Pfam=Pfam_query==Pfam_target) %>% 
  group_by(singleton) %>%
  count(is_same_3d, is_same_Pfam) %>% 
  filter(!is.na(is_same_3d), !is.na(is_same_Pfam)) %>%
  ggplot(aes(is_same_3d,is_same_Pfam,fill=n)) +
  geom_tile(color="black") +
  geom_text(aes(label=n), color="white") +
  facet_grid(singleton~.) + 
  scale_fill_gradientn(colours = wespal) + 
  coord_cartesian(expand = 0) +
  theme(legend.position = "none")
```

If we want to include multidomain proteins we can see the % of overlap between Pfam and 3D domains.

```{r multi, fig.width=10, fig.height=8}
df_multi <- df %>% 
    # head(n = 100000) %>%
  left_join(CATH_df, by=c("query"="target")) %>% 
  filter(!is.na(Pfam), !is.na(Gene3D)) %>%
  left_join(CATH_df, by="target") %>% 
  rowwise() %>% 
  mutate(Gene3D.x=strsplit(Gene3D.x, ";"),
         Pfam.x=strsplit(Pfam.x, ";"),
         Gene3D.y=strsplit(Gene3D.y, ";"),
         Pfam.y=strsplit(Pfam.y, ";")) %>% 
  mutate(n_3d_common=length(intersect(Gene3D.x,Gene3D.y)),
         n_Pfam_common=length(intersect(Pfam.x,Pfam.y))) %>% 
  # filter(n_Pfam_common>0, n_3d_common>0) %>% 
  select(-c(Pfam.x,Pfam.y,Gene3D.x,Gene3D.y)) %>% 
  mutate(overlap_3d=n_3d_common/pmax(n_3d_domains.x,n_3d_domains.y),
         overlap_Pfam=n_Pfam_common/pmax(n_Pfam_domains.x,n_Pfam_domains.y)) %>% 
  select(query, target, singleton, contains("overlap"), contains("common"))

df_multi %>% 
    pivot_longer(contains("overlap")) %>% 
    ggplot( aes(value, fill=singleton)) +
    geom_histogram(binwidth = 0.1, position = "dodge") +
    facet_grid(name~.) +
    scale_x_continuous(breaks = seq(0,1,0.1)) +
    scale_y_continuous(expand = expansion(0,0)) + 
    scale_fill_manual(values=palette_singleton)
```

<!-- This is experimental. -->

<!-- ```{r graphs, fig.height=20, fig.width=15, eval=FALSE} -->
<!-- diff <- df %>%  -->
<!--   select(-contains("method")) %>% -->
<!--   left_join(CATH_df, by=c("query"="target")) %>%  -->
<!--   left_join(CATH_df, by=c("target"), suffix = c("_query", "_target")) %>%  -->
<!--       filter(query!=target, n_Pfam_domains_query==1, n_3d_domains_query==1, -->
<!--           n_Pfam_domains_target==1, n_3d_domains_target==1) %>%  -->
<!--   filter(Pfam_query != Pfam_target | Gene3D_query != Gene3D_target) %>%  -->
<!--     select(singleton, Pfam_query, Pfam_target, Gene3D_query, Gene3D_target) -->

<!-- graph_struct <- filter(diff, Gene3D_query!=Gene3D_target) %>%  -->
<!--     group_by(singleton, Gene3D_query, Gene3D_target) %>%  -->
<!--     count() %>%  -->
<!--     filter(n>30) %>%  -->
<!--     mutate(Gene3D_query=gsub(";","",Gene3D_query), -->
<!--            Gene3D_target=gsub(";","",Gene3D_target)) %>%  -->
<!--     select(Gene3D_query, Gene3D_target, n, singleton) %>%  -->
<!--     as_tbl_graph() -->

<!-- graph_pfam <- filter(diff, Pfam_query!=Pfam_target) %>%  -->
<!--     group_by(singleton, Pfam_query, Pfam_target) %>%  -->
<!--     count() %>%  -->
<!--     filter(n>30) %>%  -->
<!--     mutate(Pfam_query=gsub(";","",Pfam_query), -->
<!--            Pfam_target=gsub(";","",Pfam_target)) %>%  -->
<!--     select(Pfam_query, Pfam_target, n, singleton) %>%  -->
<!--     left_join(nms_pfam,by=c("Pfam_query"="Pfam")) %>%  -->
<!--     left_join(nms_pfam,by=c("Pfam_target"="Pfam")) %>% -->
<!--     ungroup() %>%  -->
<!--     select(abbr.x, abbr.y, n, singleton) %>%  -->
<!--     as_tbl_graph() -->

<!-- graph_struct_plot <- ggraph(graph_struct, "graphopt") +  -->
<!--     geom_edge_fan(aes(width=n, color=singleton), show.legend = FALSE) +  -->
<!--     geom_node_point() +  -->
<!--     geom_node_text(aes(label = name),  colour = 'black', size=4, -->
<!--                   show.legend = FALSE, family = "serif", repel = T) + -->
<!--     # facet_edges(~singleton) +  -->
<!--     scale_edge_color_manual(values = palette_singleton) -->

<!-- graph_pfam_plot <- ggraph(graph_pfam, "stress") +  -->
<!--     geom_edge_fan(aes(width=n, color=singleton), show.legend = FALSE) +  -->
<!--     geom_node_point() +  -->
<!--     geom_node_text(aes(label = name),  colour = 'black', size=4, -->
<!--                   show.legend = FALSE, family = "serif", repel = T) + -->
<!--     # facet_edges(~singleton) +  -->
<!--     scale_edge_color_manual(values = palette_singleton) -->

<!-- graph_pfam_plot / graph_struct_plot -->
<!-- ``` -->

# TODO